---
title: "03-supervised"
author: "Timur Brevnov"
date: "2026-03-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Goal

The objective of this project is to predict **advertising revenue** using several supervised learning models.

Models compared:

- Linear Regression
- Forward Subset Selection
- Ridge Regression
- Lasso Regression
- PCR
- PLS
- Regression Trees
- Random Forest

Performance metrics:

- RMSE
- R²

---

## Load Libraries

```{r}
library(car)
library(leaps)
library(glmnet)
library(pls)
library(rpart)
library(rpart.plot)
library(randomForest)
```

---

## Load Dataset

```{r}
ads_data <- readRDS("data/cleaned/global_ads_performance_dataset_cleaned.rds")

ads_data$date <- NULL
ads_data$ROAS <- NULL

ads_data <- na.omit(ads_data)

str(ads_data)
```

---

## Train/Test Split

```{r}
set.seed(123)

n <- nrow(ads_data)

train_id <- sample(seq_len(n), size = floor(0.75 * n))

train_data <- ads_data[train_id, ]
test_data  <- ads_data[-train_id, ]

y_train <- train_data$revenue
y_test  <- test_data$revenue
```

---

## Linear Regression

```{r}
lm_full <- lm(revenue ~ ., data = train_data)

summary(lm_full)

vif(lm_full)

lm_pred <- predict(lm_full, test_data)

rmse_lm <- sqrt(mean((y_test - lm_pred)^2))

r2_lm <- 1 - sum((y_test - lm_pred)^2) /
  sum((y_test - mean(y_test))^2)
```
## Model Performance

The model achieved:

- **R² = 0.6885**
- **Adjusted R² = 0.6834**

This indicates that approximately **69% of the variation in revenue** can be explained by the predictors included in the model.

The overall model is statistically significant:

- **F-statistic = 133.3**
- **p-value < 2.2e-16**

### Important Predictors

The most significant predictors of revenue are:

- **Conversions**  
- **Ad Spend**  
- **Country: USA**

Conversions have the strongest effect on revenue, suggesting that increases in successful conversions lead to substantial increases in revenue.

### Multicollinearity

Variance Inflation Factor (VIF) was used to assess multicollinearity.

Most predictors have acceptable VIF values, indicating that multicollinearity is not severe, although some advertising metrics such as clicks, impressions, and ad spend show moderate correlation.
---

## Forward Subset Selection

```{r}
regfit.fwd <- regsubsets(revenue ~ .,
                         data = train_data,
                         method = "forward",
                         nvmax = ncol(train_data) - 1)

reg.summary <- summary(regfit.fwd)

plot(reg.summary$cp,
     type="b",
     pch=19,
     main="Forward Selection - Cp",
     xlab="Number of Variables",
     ylab="Cp")

best_subset_size <- which.min(reg.summary$cp)

coef(regfit.fwd, best_subset_size)
```
## Model performance

Forward subset selection was used to identify the most relevant predictors for explaining advertising revenue.

The model was evaluated using **Mallow's Cp statistic**, where lower values indicate a better balance between model fit and complexity.

The Cp plot shows that the **minimum Cp occurs at 3 predictors**, suggesting that a simple model with three variables provides the best trade-off between bias and variance.

The best subset model includes:

- countryUSA
- ad_spend
- conversions

Intercept = -224.77  
countryUSA = -3637.61  
ad_spend = 0.294  
conversions = 158.40  

### Interpretation

Conversions have the strongest positive effect on revenue, indicating that campaigns generating more conversions lead to higher revenue.

Ad spend also has a positive effect, suggesting that increased advertising investment contributes to revenue growth.

The negative coefficient for **countryUSA** indicates that campaigns targeting the United States generate slightly lower revenue compared to the baseline country category.
---

## Ridge Regression

```{r}
x_train <- model.matrix(revenue ~ . -1, train_data)
x_test  <- model.matrix(revenue ~ . -1, test_data)

cv.ridge <- cv.glmnet(x_train, y_train, alpha = 0)

plot(cv.ridge)
title("Ridge - CV MSE")

ridge_pred <- predict(cv.ridge,
                      s = "lambda.min",
                      newx = x_test)

rmse_ridge <- sqrt(mean((y_test - ridge_pred)^2))

r2_ridge <- 1 - sum((y_test - ridge_pred)^2) /
  sum((y_test - mean(y_test))^2)

ridge_coefs <- coef(cv.ridge, s = "lambda.min")

cat("Best lambda (Ridge):", cv.ridge$lambda.min, "\n\n")

print(ridge_coefs)

rmse_ridge <- sqrt(mean((y_test - ridge_pred)^2))
r2_ridge <- 1 - sum((y_test - ridge_pred)^2) /
  sum((y_test - mean(y_test))^2)
```

### Cross-Validation Results

The optimal penalty parameter \( \lambda \) was selected using **cross-validation**.
The cross-validation curve shows the relationship between the regularization parameter \( \lambda \) and the prediction error.

The vertical dashed line indicates the value of \( \lambda \) that minimizes the **cross-validated mean squared error (CVMSE)**.

As \( \lambda \) increases, coefficients are increasingly shrunk, reducing model variance but potentially increasing bias.

### Interpretation

Unlike Lasso, Ridge regression **does not remove variables**.  
Instead, it shrinks coefficients to stabilize the model when predictors are correlated.

Important predictors remain similar to those observed in previous models:

- **conversions**
- **ad_spend**
- **CTR**
- **clicks**

Conversions and advertising spend continue to have strong positive effects on predicted revenue.
---

## Lasso Regression

```{r}
cv.lasso <- cv.glmnet(x_train, y_train, alpha = 1)

plot(cv.lasso)
title("Lasso - CV MSE")

lasso_pred <- predict(cv.lasso,
                      s = "lambda.min",
                      newx = x_test)

rmse_lasso <- sqrt(mean((y_test - lasso_pred)^2))

r2_lasso <- 1 - sum((y_test - lasso_pred)^2) /
  sum((y_test - mean(y_test))^2)

lasso_coefs <- coef(cv.lasso, s = "lambda.min")

cat("Best lambda (Lasso):", cv.lasso$lambda.min, "\n\n")

print(lasso_coefs)

rmse_lasso <- sqrt(mean((y_test - lasso_pred)^2))
r2_lasso <- 1 - sum((y_test - lasso_pred)^2) /
  sum((y_test - mean(y_test))^2)
```
### Cross-Validation

The optimal value of the regularization parameter \( \lambda \) was selected using **cross-validation**, minimizing the prediction error.

In the cross-validation plot, the numbers shown above the curve represent the **number of active predictors** in the model for each value of \( \lambda \).

As \( \lambda \) increases, the model becomes simpler because more coefficients are shrunk to zero.

### Selected Predictors

Using the optimal \( \lambda \), the Lasso model retained only three predictors:

- **countryUSA**
- **ad_spend**
- **conversions**

All other predictors were shrunk to zero.

### Interpretation

The results suggest that **conversions and advertising spend are the primary drivers of revenue**.  
Campaigns targeting the **USA** have a negative coefficient relative to the baseline country (Australia).
---

## PCR

```{r}
pcr.fit <- pcr(revenue ~ .,
               data = train_data,
               scale = TRUE,
               validation = "CV")

msep_pcr <- RMSEP(pcr.fit)

rmse_pcr_cv <- sqrt(msep_pcr$val[1,1,-1])

best_pcr_comp <- which.min(rmse_pcr_cv)

pcr_pred <- predict(pcr.fit,
                    test_data,
                    ncomp = best_pcr_comp)

rmse_pcr <- sqrt(mean((y_test - pcr_pred)^2))

r2_pcr <- 1 - sum((y_test - pcr_pred)^2) /
  sum((y_test - mean(y_test))^2)

summary(pcr.fit)
validationplot(pcr.fit, val.type="MSEP")
abline(v = best_pcr_comp, col="red", lwd=2)
```
## Principal Component Regression (PCR)

Principal Component Regression (PCR) is a dimensionality reduction technique that combines **Principal Component Analysis (PCA)** with linear regression.

Instead of using the original predictors directly, PCR first transforms them into a set of **principal components**, which are uncorrelated linear combinations of the original variables.

These components are then used as predictors in the regression model.

### Cross-Validation

The optimal number of principal components was selected using **cross-validation**.

The plot shows the **Mean Squared Error of Prediction (MSEP)** as a function of the number of components.

As the number of components increases, the prediction error initially decreases because more information from the predictors is included in the model.

Based on the cross-validation results, the lowest prediction error is achieved when **22 components** are used.

### Interpretation

This indicates that most of the principal components contain useful predictive information, so the model performs best when using nearly all available components.

---

## PLS

```{r}
pls.fit <- plsr(revenue ~ .,
                data = train_data,
                scale = TRUE,
                validation = "CV")

msep_pls <- RMSEP(pls.fit)

rmse_pls_cv <- sqrt(msep_pls$val[1,1,-1])

best_pls_comp <- which.min(rmse_pls_cv)

pls_pred <- predict(pls.fit,
                    test_data,
                    ncomp = best_pls_comp)

rmse_pls <- sqrt(mean((y_test - pls_pred)^2))

r2_pls <- 1 - sum((y_test - pls_pred)^2) /
  sum((y_test - mean(y_test))^2)

summary(pls.fit)

validationplot(pls.fit,
               val.type="MSEP",
               main="PLS - CV MSEP")
abline(v = best_pls_comp, col="red", lwd=2)
```
## Partial Least Squares (PLS)

Partial Least Squares (PLS) regression is a dimensionality reduction technique similar to PCR, but it constructs components using information from both the predictors and the response variable.

Instead of creating components that only maximize the variance of the predictors (as in PCA), PLS creates components that **maximize the covariance between predictors and the response variable**.

This often leads to better predictive performance compared to PCR.

### Cross-Validation

The optimal number of components was selected using **cross-validation**.

The plot shows the **Mean Squared Error of Prediction (MSEP)** as a function of the number of components.

The vertical red line indicates the number of components that minimizes the cross-validated prediction error.

Based on the cross-validation results, the optimal model uses **9 components**.

### Interpretation

The prediction error decreases rapidly when the first components are added, indicating that they capture the most important predictive information.

After around **9 components**, the error stabilizes and additional components provide little improvement.

This suggests that the relationship between predictors and revenue can be effectively captured using a relatively small number of components.
---

## Regression Tree

```{r}
tree_model <- rpart(revenue ~ .,
                    data = train_data,
                    method = "anova",
                    control = rpart.control(cp = 0,
                                            minsplit = 5,
                                            xval = 10))

plotcp(tree_model)

best_cp <- tree_model$cptable[
  which.min(tree_model$cptable[,"xerror"]),
  "CP"]

pruned_tree <- prune(tree_model, cp = best_cp)

rpart.plot(pruned_tree)

tree_pred <- predict(pruned_tree, test_data)

rmse_tree <- sqrt(mean((y_test - tree_pred)^2))

r2_tree <- 1 - sum((y_test - tree_pred)^2) /
  sum((y_test - mean(y_test))^2)

best_cp
length(unique(pruned_tree$where))
```
## Tree Pruning

The complexity of the tree was controlled using the **complexity parameter (cp)**.

The plot shows the **cross-validated relative error** for different tree sizes.  
The optimal tree size is selected by choosing the value of \(cp\) that minimizes the cross-validated error.

Based on the cross-validation results, the selected value was:

cp = **0.00966**

After pruning the tree using this value, the final regression tree contains **5 terminal nodes**.

Pruning reduces the size of the tree and helps prevent **overfitting** while maintaining good predictive performance.

### Interpretation of the Tree

The regression tree indicates that **conversions** is the most important variable for predicting revenue.

The first split occurs at: **fewer than 221 conversions**

This means that the number of conversions is the primary factor determining revenue levels.

Further splits also depend on the number of conversions, indicating that higher numbers of conversions lead to higher predicted revenue.

For example:
campaigns with **fewer than 86 conversions** produce the lowest predicted revenue (~7000)
campaigns with **more than 704 conversions** produce the highest predicted revenue (~161000)
---

## Random Forest

```{r}
rf_model <- randomForest(revenue ~ .,
                         data = train_data,
                         ntree = 500,
                         mtry = floor(sqrt(ncol(train_data) - 1)),
                         importance = TRUE)

varImpPlot(rf_model)

plot(sqrt(rf_model$mse),
     type="l",
     main="Random Forest - OOB RMSE",
     xlab="Number of Trees",
     ylab="OOB RMSE")

rf_pred <- predict(rf_model, test_data)

rmse_rf <- sqrt(mean((y_test - rf_pred)^2))

r2_rf <- 1 - sum((y_test - rf_pred)^2) /
  sum((y_test - mean(y_test))^2)
```
## Random Forest

Random Forest is an ensemble learning method that builds a large number of decision trees and combines their predictions to improve accuracy and reduce overfitting.

### Number of Trees

The Out-of-Bag (OOB) error plot shows how the prediction error changes as the number of trees increases.

Initially, the OOB RMSE decreases rapidly as more trees are added.  
After approximately **200 trees**, the error stabilizes indicating that additional trees provide only minor improvements.

### Variable Importance

The variable importance plot ranks predictors based on their contribution to the model.

The most important variable is **conversions**, followed by **clicks**, **ad_spend**, and **impressions**.

This confirms the results obtained from earlier models, where **conversions consistently appeared as the strongest predictor of advertising revenue**.
---

## Model Comparison

```{r}
final_results <- data.frame(
  Model = c("Linear",
            "Forward Subset",
            "Ridge",
            "Lasso",
            "PCR",
            "PLS",
            "Pruned Tree",
            "Random Forest"),

  RMSE = c(rmse_lm,
           rmse_subset,
           rmse_ridge,
           rmse_lasso,
           rmse_pcr,
           rmse_pls,
           rmse_tree,
           rmse_rf),

  R2 = c(r2_lm,
         r2_subset,
         r2_ridge,
         r2_lasso,
         r2_pcr,
         r2_pls,
         r2_tree,
         r2_rf)
)

final_results <- final_results[order(final_results$RMSE), ]

final_results
```
## Model Comparison

We compared all models using **RMSE** and **R²** on the test data.

- **RMSE** measures prediction error (lower is better).
- **R²** shows how much of the variation in revenue the model explains (higher is better).

### Results

The **Forward Subset Selection** model performed best, with the lowest RMSE and the highest R².

**Lasso regression** also performed very well and gave very similar results.

The **Linear, PCR, and PLS** models produced similar performance.

The **Ridge regression** model performed slightly worse.

The **Pruned Tree and Random Forest** models had the highest prediction errors, meaning they were less accurate for this dataset.

---

## Conclusion

Overall, the results suggest that **simple linear models work best for predicting advertising revenue in this dataset**.

Across most models, **conversions** appears to be the most important predictor of revenue.