---
title: "Unsupervised Machine Learning: Advertising Campaign Segmentation"
author: "Your Name"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    toc: yes
    toc_float: yes
    number_sections: yes
---

```{r setup, include=FALSE}
# General settings for the document to keep it clean
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")

# Load required libraries
library(tidyverse)
library(factoextra)
library(cluster)

# Load the cleaned dataset
ads_data <- readRDS("data/cleaned/global_ads_performance_dataset_cleaned.rds")
```

# Executive Summary

The primary objective of this project is to apply **Unsupervised Machine Learning** techniques to analyze 1,000 global advertising campaigns. By moving beyond simple averages, we aimed to discover hidden, data-driven patterns in campaign performance.

We focused on two critical business dimensions:
* **Scale Metrics:** Ad Spend, Revenue, Conversions.
* **Efficiency Metrics:** Click-Through Rate (CTR), Cost Per Acquisition (CPA), and Return on Ad Spend (ROAS).

Through Principal Component Analysis (PCA) and K-Means Clustering, we successfully segmented the entire portfolio into **four distinct strategic profiles**, revealing that platform choice heavily dictates risk, while campaign performance remains mostly industry-neutral.

---

# Methodology: Data Preparation & Dimensionality Reduction

To ensure our clustering algorithm worked accurately, we first isolated the relevant numeric metrics and applied a logarithmic transformation (`log1p`) to scale-dependent variables (Spend, Revenue, Conversions). This prevented massive budgets from skewing the analysis.

```{r data_prep}
# Feature selection and transformation
data_clust <- ads_data %>%
  select(ad_spend, revenue, conversions, CTR, CPA, ROAS) %>%
  mutate(
    ad_spend = log1p(ad_spend),
    revenue = log1p(revenue),
    conversions = log1p(conversions)
  )

# Principal Component Analysis (PCA)
pca_res <- prcomp(data_clust, scale. = TRUE)
```

By applying **Principal Component Analysis (PCA)**, we compressed our 6 complex variables into 2 main dimensions. The first two Principal Components explain nearly **80% of the total variance**, meaning we effectively simplified the data without losing its core business meaning. 

---

# Campaign Segmentation: K-Means Clustering

Using the Silhouette method and the Gap Statistic, we determined mathematically that **4 clusters** provide the optimal segmentation for this dataset. 

```{r kmeans_clustering}
# Retain the first two components
pc_scores <- pca_res$x[, 1:2]

# Run K-Means with 4 centers
set.seed(123)
k4 <- kmeans(pc_scores, centers = 4, nstart = 50)

# Attach cluster labels to the original dataset
ads_data$cluster <- as.factor(k4$cluster)
```

## The 4 Strategic Profiles

We categorized the clusters using standard portfolio management terminology to clearly define the necessary action items for each group.

### 🌟 Cluster 3: The "Stars" (Max Volume, High Return)
* **Average Ad Spend:** €11,539
* **Average Revenue:** €63,058
* **Average ROAS:** 6.30
* **Business Insight:** These are the dominant campaigns in the portfolio. They manage the largest budgets while maintaining an excellent, highly profitable return.
* **Strategic Action:** "Feed the winners." Remove budget caps and allow these campaigns to scale as much as possible until efficiency begins to drop.

### 💎 Cluster 4: The "Hidden Gems" (Extreme Efficiency)
* **Average Ad Spend:** €2,916
* **Average Revenue:** €32,082
* **Average ROAS:** 12.2 (The highest in the dataset)
* **Business Insight:** These campaigns achieve massive results with minimal economic effort. They are highly optimized but currently underfunded.
* **Strategic Action:** Incrementally increase their daily budgets (e.g., +15% per week) to test if they can maintain this extreme efficiency at a larger scale.

### 🛡️ Cluster 1: The "Core" (Low Risk, Stable)
* **Average Ad Spend:** €1,130
* **Average Revenue:** €3,917
* **Average ROAS:** 4.52
* **Business Insight:** These represent the baseline of the account. They are small, stable, and generate safe returns without tying up large amounts of cash.
* **Strategic Action:** Keep them active as a constant, low-risk revenue stream. No immediate optimization is required.

### 🩸 Cluster 2: The "Bleeders" (Budget Wasters)
* **Average Ad Spend:** €7,723
* **Average Revenue:** €13,627
* **Average ROAS:** 2.04 (The lowest in the dataset)
* **Average CPA:** €85.50 (Extremely high acquisition cost)
* **Business Insight:** These campaigns are actively burning profit margins. You are paying a premium to acquire customers who bring in very little revenue.
* **Strategic Action:** Immediate intervention required. Pause these campaigns immediately or conduct a full creative and targeting audit before spending another dollar.

---

# Strategic Insights: Platform and Industry Trends

By analyzing where these clusters live, we uncovered actionable rules for future media buying:

```{r platforms_industries}
# Platform distribution
platform_dist <- prop.table(table(ads_data$cluster, ads_data$platform), 2)

# Industry distribution
industry_dist <- prop.table(table(ads_data$cluster, ads_data$industry), 2)
```

### 1. The Risk Profile of Platforms
* **TikTok Ads is the Growth Engine:** A massive **42.4%** of all campaigns run on TikTok end up in the "Stars" group. It is the best platform for scaling volume.
* **Google Ads is High-Risk:** Conversely, **40.4%** of Google Ads campaigns fall into the "Bleeders" category. Google requires much stricter budget controls and rigorous optimization to avoid wasted spend.
* **Meta Ads is the Efficiency King:** Meta dominates Cluster 4 (Hidden Gems), making it the ideal platform for testing highly efficient, low-budget campaigns.

### 2. The "Industry Neutrality" Concept
When looking at the distribution across E-commerce, EdTech, Fintech, Healthcare, and SaaS, the cluster proportions remain remarkably consistent. The percentage of "Stars" consistently hovers between 20% and 31% regardless of the sector. 
* **Conclusion:** Business success in this dataset is **not determined by the product you sell**, but by how the campaign is managed and the platform it is deployed on.

---

# Final Validation

To ensure our 4-cluster structure was not a random output of the K-Means algorithm, we cross-validated the results using **Hierarchical Clustering**. 

```{r hierarchical_validation, fig.cap="Dendrogram confirming the 4-cluster structure"}
data_scaled <- scale(data_clust)
dist_matrix <- dist(data_scaled)
hc_complete <- hclust(dist_matrix, method = "complete")

plot(hc_complete, main = "Dendrogram - Cluster Structure Verification", xlab = "", sub = "", labels = FALSE)
rect.hclust(hc_complete, k = 4, border = 2:5)
```

The resulting dendrogram visibly confirmed the presence of 4 distinct pillars within our data matrix, proving that our strategic segmentation is mathematically robust and reliable.