---
title: "Unsupervised Campaign Segmentation"
author: "Stela Mile"
date: "`r Sys.Date()`"
output: 
  html_document
---

```{r setup, include=FALSE}
# Chunk settings: echo=TRUE shows the code, warnings/messages are hidden for a clean report
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 6)
library(tidyverse)
library(factoextra)
library(cluster)

# Load data - Ensure your working directory is correctly set to the project root
ads_data <- readRDS("../data/cleaned/global_ads_performance_dataset_cleaned.rds")
```

# Executive Summary
This report details the application of unsupervised machine learning to segment 1,000 advertising campaigns. By moving beyond descriptive statistics, we utilize **Principal Component Analysis (PCA)** and **K-Means Clustering** to identify latent performance structures. The results reveal four distinct campaign profiles based on the interplay between financial scale and operational efficiency.

---

# Data Preparation & Feature Selection
We focus on six key performance indicators (KPIs) categorized into two dimensions:
* **Scale Metrics:** `ad_spend`, `revenue`, `conversions`.
* **Efficiency Metrics:** `CTR`, `CPA`, `ROAS`.

### Logarithmic Transformation
To address the heavy-tailed distributions typical of volume-based advertising data (where a few campaigns spend significantly more than others), we applied a `log1p` transformation. This stabilizes the variance and ensures that high-scale campaigns do not disproportionately dominate the clustering logic.

```{r feature_selection}
# Select metrics and apply transformation
data_clust <- ads_data %>%
  select(ad_spend, revenue, conversions, CTR, CPA, ROAS) %>%
  mutate(across(c(ad_spend, revenue, conversions), log1p))
```

---

# Principal Component Analysis (PCA)
We performed PCA to reduce the dimensionality of the dataset while retaining the maximum amount of information (variance).

### Variance Explanation
The Scree Plot shows that the first two components (PC1 and PC2) capture approximately **78% of the total variability**. This confirms that we can effectively analyze our campaigns using just these two dimensions.

```{r pca_scree}
pca_res <- prcomp(data_clust, scale. = TRUE)
fviz_eig(pca_res, addlabels = TRUE, ylim = c(0, 100)) + 
  ggtitle("Scree Plot: Explained Variance by Component")
```

### Variable Contributions
* **PC1 (Scale):** Dominated by `ad_spend`, `revenue`, and `conversions`.
* **PC2 (Efficiency):** Primarily driven by `ROAS`, `CPA`, and `CTR`.



```{r pca_var}
fviz_pca_var(pca_res, col.var = "contrib", repel = TRUE) +
  ggtitle("Variables Contribution to Principal Components")
```

---

# K-Means Clustering
Based on the **Silhouette** and **Gap Statistic** methods, we determined that **k = 4** is the optimal number of clusters for this dataset.

```{r kmeans_optimization, echo=FALSE}
pc_scores <- pca_res$x[, 1:2]
fviz_nbclust(pc_scores, kmeans, method = "silhouette") + 
  ggtitle("Optimal Number of Clusters (Silhouette Method)")
```

### Cluster Visualization
Using a robust initialization (`nstart = 50`), we mapped the 1,000 campaigns into four segments.



```{r kmeans_viz}
set.seed(123)
k4 <- kmeans(pc_scores, centers = 4, nstart = 50)
fviz_cluster(k4, data = pc_scores) + 
  theme_minimal() + 
  ggtitle("Campaign Segmentation Map")
```

---

# Cluster Profiling & Insights
By merging the cluster assignments back to the original data, we can profile the average performance of each segment.

```{r profiling}
ads_data$cluster <- as.factor(k4$cluster)

cluster_summary <- ads_data %>%
  group_by(cluster) %>%
  summarise(across(c(ad_spend, revenue, conversions, CTR, CPA, ROAS), mean), .groups = "drop")

knitr::kable(cluster_summary, digits = 2, caption = "Average Performance Metrics by Cluster")
```

### Strategic Segments:
1.  **Cluster 3 (Stars):** High-spend, high-revenue campaigns. These are the primary revenue drivers.
2.  **Cluster 4 (Hidden Gems):** Low spend but extremely high ROAS (12.2). These are highly efficient and prime candidates for budget scaling.
3.  **Cluster 2 (Bleeders):** High spend but very low ROAS and high CPA. These represent inefficient capital allocation.
4.  **Cluster 1 (Core):** Low-budget, stable campaigns that provide consistent baseline performance.

### Distribution Analysis
The platform analysis indicates that **TikTok Ads** frequently produces "Stars" (Cluster 3), while **Google Ads** shows a higher risk of "Bleeders" (Cluster 2). Industry distribution remains relatively uniform across clusters, suggesting that performance is driven by management and platform rather than the industry sector.

---

# Hierarchical Validation
To verify the robustness of our segmentation, we performed **Hierarchical Clustering** using complete linkage.



```{r hierarchical}
data_scaled <- scale(data_clust)
dist_matrix <- dist(data_scaled)
hc_complete <- hclust(dist_matrix, method = "complete")

plot(hc_complete, main = "Dendrogram: Hierarchical Validation", labels = FALSE)
rect.hclust(hc_complete, k = 4, border = 2:5)
```

The dendrogram confirms a consistent 4-pillar structure, matching the results found via K-Means.

---

# Final Remarks
* The segmentation highlights two main latent dimensions: **campaign scale** and **campaign efficiency**.
* PCA confirms that most variability is captured by these orthogonal dimensions, justifying dimensionality reduction.
* K-Means identifies four structurally distinct segments, suggesting meaningful heterogeneity across the dataset.
* The results should be treated as exploratory patterns, providing a data-driven foundation for budget reallocation and creative optimization.