---
title: "Global Ads Performance: Exploratory Data Analysis Report"
author: "Antonio Boscarelli"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    highlight: tango
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.width = 10, fig.height = 7)
library(tidyverse)
library(knitr)
library(reshape2)
library(lubridate)
library(gridExtra)
```

## 1. **Introduction and Dataset Overview**

```         
The goal of this Exploratory Data Analysis (EDA) is to understand
the underlying dynamics of global advertising campaign performance.
We begin by loading the cleaned dataset and observing the general
structure of the variables to identify data types (numeric vs.
categorical) and detect any major initial anomalies.
```

```{r cars}
# Caricamento del dataset
# Importing the dataset
data <- readRDS("data/cleaned/global_ads_performance_dataset_cleaned.rds")

# Initial statistical overview
str(data)
summary(data)
```

## 2. **Univariate Analysis**

## 2.1 **Numerical Variables Distribution**

In this section, we visualize the distribution shapes to detect skewness
and identify outliers. Understanding the distribution is crucial for
deciding whether to apply mathematical transformations before modeling.

```{r}
data |>
  select(impressions, clicks, ad_spend, revenue, conversions, CPC, CTR, ROAS)|>
  pivot_longer(cols = everything(), names_to = "metric", values_to = "value")|>
  
  ggplot(aes(x = value)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "#69b3a2", color = "white", alpha = 0.7) +
  geom_density(color = "red", size = 1) +
  facet_wrap(~metric, scales = "free") +
  theme_minimal() +
  labs(title = "Distribution of Numeric Metrics",
       subtitle = "Red line represents the density curve",
       x = "Value", y = "Density")
```

**Results Diagnosis:**

Right-Skewed Distributions (Positively Skewed): Metrics such as
Ad_Spend, Clicks, Revenue, Conversions, ROAS, and CPC show a peak at
lower values with a "Long Tail" extending to the right. This behavior
confirms that most campaigns have modest results, while a few outliers
achieve very high values, necessitating a Log-Transformation.

Approximate Uniform Distribution: Impressions is the only metric showing
a relatively flat distribution, indicating a balanced presence of low,
medium, and high visibility campaigns.

Normal Distribution: CTR follows a classic symmetric Bell Curve centered
around the mean.

## **2.2 Categorical Variables Frequency**

We now check for class imbalance in categorical features. An imbalanced
dataset could negatively impact the performance of future predictive
models.

```{r}
# Visualizing frequencies
data |>
  select(platform, campaign_type, industry, country) |>
  pivot_longer(cols = everything(), names_to = "dimension", values_to = "category") |> 
  count(dimension, category) |> 
  ggplot(aes(x = reorder(category, n), y = n, fill = dimension)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~dimension, scales = "free") +
  theme_minimal() +
  labs(title = "Frequency by Category", x = "Category", y = "Count")

# Summary table with percentages
summary_table <- data |>
  select(platform, campaign_type, industry, country) |>
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Category") |>
  count(Variable, Category) |>
  group_by(Variable) |>
  mutate(Percentage = sprintf("%.1f%%", n / sum(n) * 100)) |>
  ungroup() |>
  arrange(Variable, desc(n))

kable(summary_table, caption = "Category Distribution Percentages")
```

**Interpretation:**

Platform: Shows a slight imbalance, with Google Ads being the dominant
channel and TikTok Ads having the lowest frequency (\~25%).

Industry, Country, and Campaign Type: These variables exhibit a highly
uniform distribution across all categories, indicating a perfectly
balanced dataset with no "rare classes."

## 3. **Multivariate Analysis**

## 3.1 **Correlation Matrix**

The objective here is to detect multicollinearity (redundant features)
and identify strong predictors for performance metrics like ROAS.

```{r}
num_data <- data |> select(impressions, clicks, ad_spend, revenue, conversions, CTR, CPC, ROAS)
cor_mat <- cor(num_data, method = "pearson")
melted_cor <- melt(cor_mat)

ggplot(melted_cor, aes(x=Var1, y=Var2, fill=value)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(value, 2)), color = "black", size = 3) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
                       midpoint = 0, limit = c(-1,1), name="Correlation") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title = "Correlation Heatmap", 
       subtitle = "Correlation > 0.8 indicates potential multicollinearity")
```

Technical Note: A correlation $> 0.8$ implies that two variables share
over 64% of their variance ($r^2 > 0.64$). To avoid unstable
coefficients in linear models, it is recommended to select only one
feature among highly redundant pairs.

## 3.2 **Categorical Impact on Performance (Boxplots)**

We analyze ROAS and CPA (efficiency metrics) rather than volume metrics
like Clicks or Spend, as volume metrics are biased by budget allocation.
Efficiency metrics allow for a fair comparison of quality across
different segments.

```{r}
# ROAS by Platform
plot_platform <- data |>
  filter(ROAS > 0) |> 
  ggplot(aes(x = platform, y = ROAS, fill = platform)) +
  geom_boxplot(alpha = 0.7, outlier.colour = "red", outlier.shape = 1) +
  theme_minimal() +
  labs(title = "ROAS Distribution by Platform", y = "ROAS") +
  theme(legend.position = "none")

# CPA by Industry
plot_industry <- data |>
  filter(conversions > 0) |>
  ggplot(aes(x = industry, y = CPA, fill = industry)) +
  geom_boxplot(alpha = 0.7) +
  theme_minimal() +
  labs(title = "CPA (Cost per Acquisition) by Industry", y = "CPA") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")

grid.arrange(plot_platform, plot_industry, ncol = 1)
```

**Key Insights:**

ROAS by Platform: TikTok Ads shows the highest median ROAS and the
widest interquartile range, suggesting high potential returns but with
higher variance. Google Ads appears more conservative and stable.

CPA by Industry: The distribution is remarkably uniform across all
industries, indicating that the cost to acquire a customer is consistent
regardless of the sector.

## 3.3 **Ad Spend vs. Revenue (Scatterplots)**

**Goal:** Visualize the direct relationship between Spend and Revenue
across segments.

```{r scatter-spend-revenue}
data |>
  filter(revenue > 0) |> 
  ggplot(aes(x = ad_spend, y = revenue, color = campaign_type)) +
  geom_point(alpha = 0.6, size = 2) + 
  geom_smooth(method = "lm", se = FALSE) + 
  theme_minimal() +
  labs(title = "Ad Spend vs. Revenue by Campaign Type",
       subtitle = "Slope represents efficiency (Steeper = Better ROAS)",
       x = "Ad Spend", y = "Revenue")
```

**Insights:** 1. **Linearity:** Strong positive linear relationship
between Spend and Revenue across all segments. 2. **Efficiency:** Slopes
are similar, but **Display** and **Search** appear slightly steeper
(more efficient). The high overlap suggests that segmentation might
require more complex variable combinations.

## 5. **Outliers and Data Quality**

## 5.1. **Outlier Detection(quality metrics)**

```         
We focus on efficiency metrics (CPC, CPM) for outlier detection
because extreme values here often indicate technical errors,
inefficiencies, or anomalies.
```

```{r}
data2 <- data |> mutate(CPM = (ad_spend / impressions) * 1000)

data2 |>
  select(CPC, CPM) |>
  pivot_longer(cols = everything(), names_to = "metric", values_to = "value") |>
  ggplot(aes(x = metric, y = value, fill = metric)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 8, outlier.size = 2) +
  theme_minimal() +
  facet_wrap(~metric, scales = "free") +
  labs(title = "Outlier Detection in Cost Metrics", 
       subtitle = "Red stars indicate statistical outliers (1.5 * IQR)", y = "Value") +
  theme(legend.position = "none")
```

Visual Inspection:Both metrics exhibit a heavy tail with significant
outliers. In particular, CPM shows values reaching $>200$ while the
median is $\sim50$. This confirms the need to scale or transform the
data to mitigate the impact of these extreme values.

## 5.2. **Zero-Revenue Analysis (Budget Waste)**

**Goal:** Quantify campaigns that spent money but generated zero
revenue.

```{r waste-analysis}
waste_check <- data |>
  mutate(performance_status = ifelse(revenue == 0, "Zero Revenue (Waste)", "Profitable")) |>
  count(performance_status) |>
  mutate(percentage = sprintf("%.1f%%", n / sum(n) * 100))

kable(waste_check, caption = "Budget Waste Check")
```

**Result:** 100% of campaigns are profitable according to the dataset.

## 6. **Log-Transformation Validation**

```         
Following the skewness found in section 2.1, we validate the
effectiveness of the Log-Transformation in normalizing the data.
```

```{r}
p1_orig <- ggplot(data, aes(x = ad_spend)) + 
  geom_histogram(fill = "#e74c3c", bins = 30, color = "white", alpha = 0.8) +
  labs(title = "Original Ad Spend", subtitle = "Right Skewed", x = "Ad Spend ($)") + theme_minimal()

p1_log <- ggplot(data, aes(x = log1p(ad_spend))) + 
  geom_histogram(fill = "#2ecc71", bins = 30, color = "white", alpha = 0.8) +
  labs(title = "Log-Transformed Ad Spend", subtitle = "Normalized", x = "Log(Ad Spend + 1)") + theme_minimal()

# 6.2 CPC Comparison
p2_orig <- ggplot(data, aes(x = CPC)) + 
  geom_histogram(fill = "#e74c3c", bins = 30, color = "white", alpha = 0.8) +
  labs(title = "Original CPC", subtitle = "Extreme Outliers", x = "CPC ($)") + theme_minimal()

p2_log <- ggplot(data, aes(x = log1p(CPC))) + 
  geom_histogram(fill = "#2ecc71", bins = 30, color = "white", alpha = 0.8) +
  labs(title = "Log-Transformed CPC", subtitle = "Compressed Outliers", x = "Log(CPC + 1)") + theme_minimal()

# Display full grid
grid.arrange(p1_orig, p1_log, p2_orig, p2_log, ncol = 2)
```

Final Transformation ValidationThe comparison plots confirm that the
Log-Transformation ($log1p$) is highly effective for preparing this
dataset for advanced modeling:Ad Spend (Top Row): The transformation
successfully handles the heavy right-skewness. The distribution shifts
from being clustered on the far left to a near-perfect Gaussian curve,
which is ideal for regression tasks.CPC (Bottom Row): In this case, the
transformation successfully compresses the variance. It pulls the
extreme outliers from the right tail significantly closer to the center
of the distribution.
